\vspace{-1.3cm} \textmd{\normalsize{High Performance Computing Project}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration

### 2016-03-29 paper/paper.tex
paper...

%\pagestyle{fancy}
%\fancyhead[L]{Seminar \glqq Optimale Steuerung\grqq\\ Prof. Brokate, Prof. Kuttler} %Kopfzeile links
%\fancyhead[C]{} %zentrierte Kopfzeile
%\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
\title{\vspace{-1.3cm} \textmd{\normalsize{Project description}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
\date{}
\author{Manuel Demmeler}

@@

Then in a first level of parallelization, this function can be applied to the \(q_i \in [q, \tilde q]_h \) to check, if \(q\) and \(\tilde q\) are connected.

Furthermore, a second level of parallelization could be achieved by sampling and connecting several new points \(\tilde q_i\) parallely. This is expected to be more difficult, because all new nodes have to be connected to the old graph and to each other. Therefore it has to be evaluated, if a parallel implementation stays efficient here.

If there remains time, a next idea could be to execute the algorithm on different discretization levels \(h\). This means for example to compute a way \(q_1^0, ..., q_n^0\) which is connected with a coarse stepsize \(h_0\), then to decrease the stepsize to \(0<h_1<h_0\) and take the path nodes \(q_i^0\) as starting nodes for the next execution of the algorithm, which gives then an \(h_1\)-connected way \(q_1^1, ... , q_n^1\), etc.

\section{Structure of the PRM Solver}

@@

\section{Implementation of an Indicator Function for Robot Arms}

In the project, the probabilistic roadmap algorithm is applied  to create collisionfree trajectories of a robot arm.
For a robot with given joint angles \(q_i\) one can calculate straightforwardly all positions and orientations  of the single parts of the robot.
For this, coordinate frames according to the Denavit-Hartenberg convention are used.
The coordinate frames of the single parts are described by transformation matrices 
$T_i \in R^{4x4}$, which transform points of the body frame into the world frame. They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot.

The geometry of the robot parts is defined by a set of convex polytopes, each of them belonging to one of the robots coordinate frame. Applying the corresponding $T_i$, its vertices can be transformed into the world frame. 
Environment obstacles are also defined by convex polytopes, in the world frame.

Now the movement of the robot can be described by a curve $q(t)$ in the state space. It is collisionfree, if at no time, any two polytopes of the robot intersect with each other or with the environment after transformed into the world frame.
To apply the probabilistic roadmap algorithm, we express this by the indicator function

\begin{equation*}
	I(q)=
	\begin{cases}
		0,	& \text{ for no collision for all given pairs of polytopes in wold frame } \\
		1,	& \text{ if at least one collision occurs }
	\end{cases}
\end{equation*}

The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread. Therefore, the kinematics and collision algorithm have been implemented as device code for the use in the kernel.

The Denavit-Hartenberg (DH) convention is a method to describe the geometry of an robots joints, 
that is, how the transformations between the robots coordinate frames look like in dependancy of its joint values $q_1,...,q_d$.
For this it is assumed that each joint is either rotational, that means an rotation around a fixed axis, 
or prismatic, a translation along an axis.
The robot consists of $d+1$ bodies each of which has an body fixed coordinate frame $B_i$, which is described by the transformation matrix $T_i$ into the world frame.
For an robot with DH parameters $a_{i-1}, \alpha_{i-1}, d_i \in \R$, $i=1,...,d$, 
the relative transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies now is defined as
\begin{equation}
	T_{i-1,i} = R_z(\theta_i) T_z(d_i) R_x(\alpha_{i-1}) R_x(a_{i-1})
\end{equation}
where $R_x, R_z$ are rotations and $T_x, T_z$ translations around the $x$ and $z$ axes.
By an proper choice of the body frames, every robot with rotational or prismatic joint can be described by DH parameters.
For the exact calculations, see \cite{robodyn}.

Transformation matrices are implemented float4 arrays with inline functions for common matrix operations. Again operators with temporaries avoided.
\begin{lstlisting}
class trafo4{
  public:
    float4 col[4];
    __host__ __device__ trafo4(){}
    __host__ __device__ trafo4(float a, float alpha, float q, float d);
    __host__ __device__ inline float4& apply(float4& u) const;
    __host__ __device__ inline float4& apply(const float4& u, float4& Tu) const;
    .
    .
    .
}
\end{lstlisting}

>>>>

%\pagestyle{fancy}
%\fancyhead[L]{Seminar \glqq Optimale Steuerung\grqq\\ Prof. Brokate, Prof. Kuttler} %Kopfzeile links
%\fancyhead[C]{} %zentrierte Kopfzeile
%\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
\title{\vspace{-1.3cm} \textmd{\normalsize{High Performance Computing Project}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
\date{}
\author{Manuel Demmeler}

@@

While the distribution of the indicator function to different CUDA threads is a first level of parallelization, 
a second level can be achieved by sampling and connecting several new nodes parallely. 
This is more difficult, because all new nodes have to be connected 
and the graph data has to be updated on all processes. 
For this, different approaches have been implemented, which are described in chapter \ref{solver}.

Before that, chapter \ref{indicator} gives more details about the robotics indicator function.

\section{Implementation of an Indicator Function for Robot Arms}\label{indicator}.

In the project, the probabilistic roadmap algorithm is applied  to create collisionfree trajectories of a robot arm.
For a robot with given joint angles \(q_i\) one can calculate straightforwardly all positions and orientations  of the single parts of the robot.
For this, coordinate frames according to the Denavit-Hartenberg convention are used.
The coordinate frames of the single parts are described by transformation matrices 
$T_i \in R^{4x4}$, which transform points from the body frame frame into the world frame. They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot.

The geometry of the robot parts is defined by a set of convex polytopes, each of them belonging to one of the robots coordinate frame. Applying the corresponding transformation $T_i$, its vertices can be transformed into the world frame. 
Environment obstacles are also defined by convex polytopes, in the world frame.

Now the movement of the robot can be described by a curve $q(t) \in \R^d$ in the configuration space. It is collisionfree, if at no time, any two polytopes of the robot intersect with each other or with the environment after being transformed into the world frame.
To apply the probabilistic roadmap algorithm, we express this by the indicator function

\begin{equation*}
	I(q)=
	\begin{cases}
		0,	& \text{ for no collision for all given pairs of polytopes } \\
		1,	& \text{ if at least one collision occurs }
	\end{cases}
\end{equation*}

The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread. Therefore, the kinematics and collision algorithm are implemented as device code for the use in the kernel.

The Denavit-Hartenberg (DH) convention is a method to describe the geometry of an robots joints, 
that is, how the transformations between the robots coordinate frames look like in dependancy of its joint values $q_1,...,q_d$.
For this it is assumed that each joint is either rotational, that means an rotation around a fixed axis, 
or prismatic, a translation along an axis.
The robot consists of $d+1$ bodies each of which has an body fixed coordinate frame $B_i$, which is described by a transformation matrix $T_i$ into the world frame.
For a robot with DH parameters $a_{i-1}, \alpha_{i-1}, d_i, \theta_i \in \R$, $i=1,...,d$, 
it is now defined, that the relative transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies are
\begin{equation}
	T_{i-1,i} = R_z(\theta_i) T_z(d_i) R_x(\alpha_{i-1}) R_x(a_{i-1})
\end{equation}
where $R_x, R_z$ are rotations and $T_x, T_z$ translations around the $x$ and $z$ axes.
From this follows, that from given DH parameters all transformations are determined.
The other way round, one can show, that by an proper choice of the body frames, every robot with rotational or prismatic joints can be described by a set of DH parameters.
For the exact calculations, see \cite{robodyn}.

Transformation matrices are implemented as float4 arrays with inline functions for common matrix operations. Again operators with temporaries were avoided, wherever possible.
\begin{lstlisting}
class trafo4{
  public:
    float4 col[4];
    __host__ __device__ trafo4(){}
    __host__ __device__ trafo4(float a, float alpha, float q, float d);
    __host__ __device__ inline float4& apply(float4& u) const;
    __host__ __device__ inline float4& apply(const float4& u, float4& Tu) const;
    .
    .
    .
}
\end{lstlisting}

\section{Structure of the PRM Solver}\label{solver}

@@

\section{Results}

- timings of different versions at same conditions

\section{Conclusion}
